CONEM2024-0025\\ ABC MULTICRITERIA CLASSIFICATION WITH ANALYTICAL HIERARCHY PROCESS: HOW TO FORCE CONSISTENCY OF PAIRWISE COMPARISONS MATRIX

### 2024-02-16 accepted/CONEM2024/FCPCOM/CONEM2024.tex
updating
work in progress - writing introduction

    \subsection{Pairwise Comparisons Matrix and Consistency Rate}

@@

>>>>

	As stated by \cite{Kiran2019}, the "production planning and control (PP\&C) ... is responsible for ensuring the availability all materials, part of assembly at the right time, at the right place, and in right quantities in order to enable the progress of operations according to the predetermined schedules at the minimum possible costs." According to the author, factories producing large volumes of standardized products need very simple production control techniques. However factories producing a variety of products and product mix also need meticulous control with complex production control techniques. Therefore, in the latter case, a prior strategic planning becomes impressible for better use of the productive capacity as well as to avoid investment losses in materials that will not be used or will not bring significant profit to the company. Companies that market a wide variety of products and product mix should always use good strategic planning as a guide for sales and PP\&C anticipation strategies and it's really important to start any strategic planning by understanding the importance all products marked.

	In this work we conducted a study with three companies from three different sectors, which market a wide variety of products, in some cases, highly personalized products. In view of the very high variety of products marketed by the companies studied, the interest of this study was to identify the main products, that is, the products that should be prioritized, so that this information could guide the strategic planning of the company. Also, we identified the need to consider several criteria for such a classification, therefore we used a multicriteria approach.

	The importance of multicriteria classification for inventory management has been widely emphasized in scientific literature and several multicriteria methods have been used, being the ABC multicriteria classification with Analytical Hierarchy Process one of the most common techniques applied for this purpose \citep{Flores1992, AltayGuvenir1998, Lolli2014, Balaji2014}. Few recent works talk about multicriteria classification applied to strategic planning \citep[\emph{e.g.},][]{BarbosaDePaula2022, Pereira2023, MarianoRibeiro2023}.

	\citet{Odu2019} emphasize that, when aplying some multicriteria method, it is important to pay particular attention to the objectivity factors of criteria weights. The author provides an important overview of different weighting methods applicable to multicriteria techniques. In the present work, we opted for the multicriteria ABC classification and adopted the Analytic Hierarchy Process \citep{Saaty1987} to assign the weights to the different criteria. It is important to note that the Analytic Hierarchy Process derives the weights of the criteria based on the eigenvector of a pairwise comparison matrix which is constructed manually according to an empirical evaluation. According to \cite{Saaty1987}, this matrix must be antisymmetric and must have a consistency rate of less than 0.1.  The autor reports that a software package supporting the AHP, called Expert Choice, for the IBM PC, was used in his work for making eigenvector calculations and guide the decision maker to improve matriz inconsistency if needed. During the execution of the work we present in this paper, we found that there is a strong difficulty in for calculation of the eigenvector, as well as for the construction of consistent pairwise comparison matrices and we had no access to the Expert Choice software package, so we developed an open source COPSolver library for Analitic Hierarch Process and multicriteria ABC classification based on a new a method for forcing the consistency of such matrices. We present this method in this paper. The Eigen library from eigen.tuxfamily.org was used for eigenvector calculation. Also we identified that the assignment of weights to ABC classification can be done using different criteria and we observed that such assignment generates a bias in the multicriteria ABC classification. Therefore we also developed a new methodology for assigning weights in a balanced way for ABC classification and we proposed a new multicriteria ABC classification method, that we aplyed on COPSolver library. We also presente both methods on this paper and an in-depth analysis of the results found by the library we developed for COPSolver.

@@

    Inventory classification problem is also a decision problem since we want to decide what is the better system and criteria for clustering the different inventory itens.
